---
title: "team.Rmd"
author: "Muhamad Imran Bin Maszeri and Dongming Li"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploration

Use ggplot2 to create the plots, and dplyr to manipulate the data frame as much as possible. You may want to rename the columns of the data frame for your convenience using rename in dplyr.

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(dplyr)
library(tidyverse)
dat <- readxl::read_xls('GSS2.xls')
dat1<-rename(dat,marital=`Marital status`, educ=`Highest year of school completed`, sex=`Respondents sex`, party=`Political party affiliation`, happy=`General happiness`, fincome=`Total family income`, age=`Age of respondent`)
dat1<-dat1%>%filter(educ!="No answer")
dat1<-dat1%>%filter(educ!="Don't know")
dat1<-dat1%>%filter(marital!="No answer")
dat1<-dat1%>%filter(party!="No answer")
dat1<-dat1%>%filter(party!="Don't know")


dat1 <- dat1 %>% mutate(educ=as.numeric(educ))
EDU<-select(dat1,educ, party)
EDU<-group_by(EDU,party)
EDU<-drop_na(EDU)
EDU<-summarise(EDU,AVEEDU=mean(educ))
EDU
```
    
    The output shows the average years of education for each political party affiliation. From the output, in general, we can see Republicans at all levels are better educated than Democrats at the same level. And individuals' education level is a little lower than these two affiliation.
    
2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

For each of the investigation below, create at least one numerical summary and one appropriate visualization. Write a paragraph to describe your findings and the variables you used.

3. How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is “Very happy” or “Pretty happy”.

```{r echo=FALSE, message=FALSE, warning=FALSE}
dat2<-dat1 %>% filter(happy %in% c("Pretty happy", "Very happy","Not too happy"))
Happiness<-select(dat2,happy,marital)
Happiness<-group_by(Happiness,marital)
Happiness<-drop_na(Happiness) 
summarise(Happiness,CountHappy=n())
Happiness%>%ggplot(aes(x=marital, fill=happy)) + geom_bar(position='dodge')
```
    
    From the data and plot we can see that most of responders in this data are married, and most people who married are happy. Also we can see that for other maritals, the total number of responders decrease but the people who are not too happy have a higher proportion. That shows unmarried marital does have a negative impact on happiness.
    
4. Does party affiliation affect the relationship you found in the last question?

5. How is marital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?

```{r echo=FALSE,message=FALSE,warning=FALSE}
dat3<-select(dat1,educ,marital)
dat3<-group_by(dat3,marital)
dat3<-drop_na(dat3)
Summary<-dat3%>%summarise(AVEedu=mean(educ))
Summary
dat3%>%ggplot(aes(educ, marital, fill=marital)) + geom_boxplot()
```

    In general, from data we can see people who never married have the highest average educarion level. This shows that the higher education level, the harder for people to find a good marriage. Because people with high education level may increase their requirement for marriage.

6. Explore two more interesting questions, and answer those questions using the GSS data.
