---
title: "team.Rmd"
author: "Muhamad Imran Bin Maszeri and Dongming Li"
date: "9/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data reading and cleaning. Use GSS and GSS2. I think we should try to combine this
```{r message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
library(dplyr)
datum <- readxl::read_xls('GSS.xls')
newdat <- datum %>%
  rename(year=`Gss year for this respondent`, 
         marital=`Marital status`, 
         educ=`Highest year of school completed`, 
         sex=`Respondents sex`, 
         partyid=`Political party affiliation`, 
         happy=`General happiness`, 
         ballot=`Ballot used for interview`, 
         id=`Respondent id number`,
         conpress=`Confidence in press`,
         bills=`Income alone is enough`,
         qlife=`R's quality of life`) %>%
  filter(!educ %in% c("No answer", "Don't know")) %>%
  filter(!marital %in% c("No answer")) %>%
  filter(!partyid %in% c("No answer", "Don't know"))

dat <- readxl::read_xls('GSS2.xls')
```

## Exploration

Use ggplot2 to create the plots, and dplyr to manipulate the data frame as much as possible. You may want to rename the columns of the data frame for your convenience using rename in dplyr.

1. Find the average years of education for the respondents with each political party affiliation. Arrange your output in a meaningful order and print. Describe any patterns you find.
```{r echo = FALSE, message=FALSE, warning=FALSE}
dat1<- select(dat, 'Highest year of school completed', 'Political party affiliation')
Indneadem<-dat1[dat1$`Political party affiliation`=='Ind,near dem',]
Indneadem<-Indneadem[!(Indneadem$`Highest year of school completed`== 'No answer'),]
Indneadem<-Indneadem[!(Indneadem$`Highest year of school completed`== "Don't know"),]
Indneadem<-Indneadem[rowSums(is.na(Indneadem))!=ncol(Indneadem),]
EDU<-as.numeric(Indneadem$`Highest year of school completed`)
AVGIndneadem<-mean(EDU)
Indnearrep<-dat1[dat1$`Political party affiliation`=='Ind,near rep',]
Indnearrep<-Indnearrep[!(Indnearrep$`Highest year of school completed`== 'No answer'),]
Indnearrep<-Indnearrep[!(Indnearrep$`Highest year of school completed`== "Don't know"),]
Indnearrep<-Indnearrep[rowSums(is.na(Indnearrep))!=ncol(Indnearrep),]
EDU2<-as.numeric(Indnearrep$`Highest year of school completed`)
AVGIndnearrep<-mean(EDU2)
INDEPENDENT<-dat1[dat1$`Political party affiliation`=='Independent',]
INDEPENDENT<-INDEPENDENT[!(INDEPENDENT$`Highest year of school completed`== 'No answer'),]
INDEPENDENT<-INDEPENDENT[!(INDEPENDENT$`Highest year of school completed`== "Don't know"),]
INDEPENDENT<-INDEPENDENT[rowSums(is.na(INDEPENDENT))!=ncol(INDEPENDENT),]
EDU3<-as.numeric(INDEPENDENT$`Highest year of school completed`)
AVGINDEPENDENT<-mean(EDU3)
Notstrdemocrat<-dat1[dat1$`Political party affiliation`=='Not str democrat',]
Notstrdemocrat<-Notstrdemocrat[!(Notstrdemocrat$`Highest year of school completed`== 'No answer'),]
Notstrdemocrat<-Notstrdemocrat[!(Notstrdemocrat$`Highest year of school completed`== "Don't know"),]
Notstrdemocrat<-Notstrdemocrat[rowSums(is.na(Notstrdemocrat))!=ncol(Notstrdemocrat),]
EDU4<-as.numeric(Notstrdemocrat$`Highest year of school completed`)
AVGNotstrdemocrat<-mean(EDU4)
Notstrrepublican<-dat1[dat1$`Political party affiliation`=='Not str republican',]
Notstrrepublican<-Notstrrepublican[!(Notstrrepublican$`Highest year of school completed`== 'No answer'),]
Notstrrepublican<-Notstrrepublican[!(Notstrrepublican$`Highest year of school completed`== "Don't know"),]
Notstrrepublican<-Notstrrepublican[rowSums(is.na(Notstrrepublican))!=ncol(Notstrrepublican),]
EDU5<-as.numeric(Notstrrepublican$`Highest year of school completed`)
AVGNotstrrepublican<-mean(EDU5)
Strongdemocrat<-dat1[dat1$`Political party affiliation`=='Strong democrat',]
Strongdemocrat<-Strongdemocrat[!(Strongdemocrat$`Highest year of school completed`== 'No answer'),]
Strongdemocrat<-Strongdemocrat[!(Strongdemocrat$`Highest year of school completed`== "Don't know"),]
Strongdemocrat<-Strongdemocrat[rowSums(is.na(Strongdemocrat))!=ncol(Strongdemocrat),]
EDU6<-as.numeric(Strongdemocrat$`Highest year of school completed`)
AVGStrongdemocrat<-mean(EDU6)
Strongrepublican<-dat1[dat1$`Political party affiliation`=='Strong republican',]
Strongrepublican<-Strongrepublican[!(Strongrepublican$`Highest year of school completed`== 'No answer'),]
Strongrepublican<-Strongrepublican[!(Strongrepublican$`Highest year of school completed`== "Don't know"),]
Strongrepublican<-Strongrepublican[rowSums(is.na(Strongrepublican))!=ncol(Strongrepublican),]
EDU7<-as.numeric(Strongrepublican$`Highest year of school completed`)
AVGStrongrepublican<-mean(EDU7)
Otherparty<-dat1[dat1$`Political party affiliation`=='Other party',]
Otherparty<-Otherparty[!(Otherparty$`Highest year of school completed`== 'No answer'),]
Otherparty<-Otherparty[!(Otherparty$`Highest year of school completed`== "Don't know"),]
Otherparty<-Otherparty[rowSums(is.na(Otherparty))!=ncol(Otherparty),]
EDU8<-as.numeric(Otherparty$`Highest year of school completed`)
AVGOtherparty<-mean(EDU8)
POP<-c(AVGIndneadem,AVGIndnearrep,AVGNotstrdemocrat,AVGNotstrrepublican,AVGStrongdemocrat,AVGStrongrepublican,AVGINDEPENDENT,AVGOtherparty)
names(POP)<-c('Independent near democrat','Independent near republican','Not strong democrat','Not strong republican','Strong democrat','Strong republican','Independent','Other')
POP
```
    
    The output shows the average years of education for each political party affiliation. From the output, in general, we can see Republicans at all levels are better educated than Democrats at the same level. And individuals' education level is a little lower than these two affiliation.
    
2. Create a single data frame containing records for the better educated respondents in each political party affiliation. A “better educated” respondent is someone who has strictly more years of education than the median in the same party affiliation. Print the structure of the data frame.

```{r message=FALSE, warning=FALSE}
df1 <- newdat %>%
  mutate(educ=as.numeric(educ)) %>%
  group_by(partyid) %>%
  mutate(median_educ=median(educ, na.rm = TRUE)) %>%
  filter(educ>median_educ) %>%
  ungroup()

str(df1)
```

For each of the investigation below, create at least one numerical summary and one appropriate visualization. Write a paragraph to describe your findings and the variables you used.

3. How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is “Very happy” or “Pretty happy”.
```{r echo=FALSE}

```

4. Does party affiliation affect the relationship you found in the last question?

```{r message=FALSE, warning=FALSE}
newdat %>%
  select(happy, marital, partyid) %>%
  filter(happy %in% c("Pretty happy", "Very happy")) %>%
  drop_na() %>%
  group_by(marital, partyid) %>%
  summarise(nHappy=n()) %>%
  arrange(desc(nHappy))

newdat %>%
  select(happy, marital, partyid) %>%
  filter(happy %in% c("Not too happy", "Pretty happy", "Very happy")) %>%
  drop_na() %>%
  ggplot(aes(x=marital, fill=happy)) + geom_bar(position='stack') + labs(fill="Happiness") + facet_wrap(~partyid) + coord_flip()
```

Answer: Not really, all of the party affiliations have a very high happiness for married respondents. Happy, Married respondents are the highest and have an affilliation for not strong democrat.

5. How is marital status related to the education of a respondent? What are possible reasons accounting for the patterns you see?

6. Explore two more interesting questions, and answer those questions using the GSS data.

Question 1: How do party affilliation affects their confidence towards the press, with each sex?
```{r message=FALSE, warning=FALSE}
newdat %>%
  select(sex, conpress, partyid) %>%
  filter(!conpress %in% c("Not applicable", "No answer")) %>%
  drop_na() %>%
  ggplot(aes(x=partyid, fill=conpress)) + geom_bar(position='dodge') + labs(title="Confidence in press with Party Affilliation", x="Party Affilliation", fill="Press Confidence") + facet_wrap(~sex) + coord_flip()
```

Answer: We can see that almost all of the party affilliation has only some confidence towards the press, but for "Strong Republican" respondents, they are really interesting. Male strong republicans are mostly do not feel confident with the press, while female strong republicans are very close between believing only some press and believing hardly any press. This means that strong republican respondents don't really trust the current news.
